---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
The next code download and unzip the csv file data into temporal folder:

```{r download, include=TRUE}
temp <- tempdir()
urlfile<-"https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"

if(!file.exists(paste(temp,"data.zip",sep = "/"))){
      #dir.create(paste(temp,"data",sep = "/"))
      download.file(urlfile, paste(temp,"data.zip",sep="/"))
      }
unzip(paste(temp,"data.zip",sep="/"), exdir = paste(temp,"data",sep = "/"))
```

Now, we read the csv file in R and coercing dates
```{r loaddata, include=TRUE}
activity<-read.csv(paste(temp,"data","activity.csv",sep="/"))
activity$date<-as.Date(activity$date, format="%Y-%m-%d")
```

## What is mean total number of steps taken per day?
### 1. Total number of steps taken per day
```{r totalSteps, message=FALSE, warning=FALSE}
require(data.table)
StepsPerDay<-data.table(activity)[, list(StepsPerDay=sum(steps, na.rm = T)), by=c("date")]
head(StepsPerDay,10)
```

### 2. Histogram of the total number of steps taken each day
```{r histogram, message=FALSE, warning=FALSE}
hist(StepsPerDay$StepsPerDay/1000, main = "Steps taken each day", 
     xlab = "Number of steps (Thousands)", ylab="Frecuency of steps", 
     col="deepskyblue3", border = "deepskyblue4", breaks = 20)
abline(v=mean(StepsPerDay$StepsPerDay/1000), lty=3, lwd=2, col="blue")
legend(legend="Mean", "topright", lty=3, lwd=2, col="blue",bty = "n")
```

### 3. The mean and median of the total number of steps taken per day
```{r mean, message=FALSE, warning=FALSE}
mean(StepsPerDay$StepsPerDay)
median(StepsPerDay$StepsPerDay)
```

## What is the average daily activity pattern?
### 1. Time series plot of the 5-minute interval and the average number of steps taken averaged across all days.
```{r tseries, message=FALSE, warning=FALSE}
PatternMean<-data.table(activity)[, list(Mean=mean(steps, na.rm = T)), by=c("interval")]
head(PatternMean,10)
plot(PatternMean$interval, PatternMean$Mean, type = "l", col="deepskyblue4",
     main = "Pattern of steps by interval", xlab="Interval", ylab = "Average steps")
```

### 2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r max, message=FALSE, warning=FALSE}
maxInterval<-PatternMean$interval[which.max(PatternMean$Mean)]
```
The interval with the maximun average steps is the `r max(maxInterval)`, with `r max(round(PatternMean$Mean[which.max(PatternMean$Mean)],2))` average steps.

## Imputing missing values
### 1. Total number of missing values in the dataset
```{r missing, message=FALSE, warning=FALSE}
require(dplyr)
dim(filter(activity, is.na(steps)))[1]
```

### 2. Strategy for filling in all of the missing values in the dataset.
We use the mean for that 5-minute interval:
```{r strategy}
fillData<-merge(activity,PatternMean, by="interval", all.x = T)
fillData<-mutate(fillData, steps=ifelse(is.na(steps),Mean,steps)) 
head(fillData, 10)
```

### 3. New dataset that is equal to the original dataset but with the missing data filled in
```{R NewDB}
fillData<-select(fillData, steps, date, interval)
fillData<-fillData[order(fillData$date),]
head(fillData, 10)
```

### 4. Histogram and impact of missing values
```{r hist2}
StepsPerDay2<-data.table(fillData)[, list(StepsPerDay=sum(steps, na.rm = T)), by=c("date")]
hist(StepsPerDay2$StepsPerDay/1000, main = "Steps taken each day", 
     xlab = "Number of steps (Thousands)", ylab="Frecuency of steps", 
     col="deepskyblue3", border = "deepskyblue4", breaks = 20)
abline(v=mean(StepsPerDay2$StepsPerDay/1000), lty=3, lwd=2, col="blue")
abline(v=median(StepsPerDay2$StepsPerDay/1000), lty=3, lwd=2, col="red")
legend(legend="Mean and Median", "topleft", lty=3, lwd=3, col="red",bty = "n")
```


## Are there differences in activity patterns between weekdays and weekends?
### 1. New factor variable in the dataset with two levels – “weekday” and “weekend”
```{r TypeDay}
require(dplyr, lubridate)
WeekDay<-function(date){
      if(wday(date) %in% c(1,7)) TypeDay<-"weekend"
      else {
            TypeDay<-"weekday"
      }
      return(TypeDay)
}

fillData<-mutate(fillData, TypeDay=as.factor(sapply(date, WeekDay)))
levels(fillData$TypeDay)
table(fillData$TypeDay)
```

### 2. Panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis)

```{r PanelPlot, message=FALSE, warning=FALSE}
require(ggplot2)
DailyPattern<-data.table(fillData)[, list(Mean=mean(steps, na.rm = T)), by=c("interval","TypeDay")]
head(DailyPattern,10)
p <- ggplot(DailyPattern, aes(interval, Mean)) + geom_line()
p + facet_grid(TypeDay ~ .)

```



